\input{header}

%% thesis

\thesislang{en}                   % The language of the thesis
% The title of the thesis
\thesistitle{Efficient exhaustive generation of small multiway cuts}
\thesissubtitle{Bachelor Thesis}  % The type of the thesis
\thesisstudent{Ondřej Slámečka}   % Your name
\thesiswoman{false}               % Your gender
\thesisfaculty{fi}                % Your faculty
\thesisyear{Spring \the\year}     % The academic term of your thesis defense
\thesisadvisor{prof. RNDr. Petr Hliněný, Ph.D.}   % Your advisor

\begin{document}
  \FrontMatter                    % The front matter
    \ThesisTitlePage                % The title page
    \begin{ThesisDeclaration}       % The declaration
      \DeclarationText
      \AdvisorName
    \end{ThesisDeclaration}
    \begin{ThesisThanks}            % The acknowledgements (optional)
      I would like to thank my supervisor prof. RNDr. Petr Hliněný, Ph.D. who suggested this interesting topic to me and guided me throughout my work.


    \end{ThesisThanks}
    \begin{ThesisAbstract}          % The abstract

Given a large but sparse graph the problem is to generate all minimal multiway edge-cuts up to a certain number of edges. We present a detailed description of a newly proposed "circuit-cocircuit" algorithm and its modification for almost canonical generation of cuts, based on the technique of generation by canonical construction path. We provide an implementation of the algorithm in a computer program and evaluate its various features to demonstrate practical feasibility of the algorithm. In the evaluation a special attention is paid to real-world road networks.

	\end{ThesisAbstract}
    \begin{ThesisKeyWords}          % The keywords
      graph theory, matroid theory, graph cuts, graph algorithms, algorithm implementation
    \end{ThesisKeyWords}
    \tableofcontents                % The table of contents
%   \listoftables                   % The list of tables (optional)
%   \listoffigures                  % The list of figures (optional)

\MainMatter                     % The main matter

\include{ch_introduction}
\include{ch_definitions}
\include{ch_algorithm}
\include{ch_canonical}
\include{ch_implementation}
\include{ch_evaluation}
\include{ch_conclusion}

	% Bibliography goes here

\printbibliography

    % Index goes here (optional)
\end{document}
